TR Invariant T.I.

### 2017-04-26 Notes of various papers/TR invariant topological insulators T. Hughes. Dissertation/note.tex
[major] Try to get the Edge modes and Chern number of Lattice Dirac Ham in Hughes's dissertation.

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physicist-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}
\makenomenclature % For an index of symbols.

@@

This amounts to removing the $A$ in top right corner and $C$ in bottom
left corner of eq.\ref{eq:disc-eigeneq}. 
\footnote{The calculation can be found
in file \texttt{Lattice Dirac Model (2+1)-d-4-Infinite Stripe
(N=3).nb}, and file \texttt{Lattice Dirac Model (2+1)-d-4-Infinite
Stripe (N=20).nb}.}
In this case, the crossing bands exists at slightly different value of
$m$, also, there are few sharp transitions from a closed gap state, to an
open gap state. For example, when $N=3$, the band crossing happens at
two range of $m$: from $0.65$ to $1.35$ (shown in
fig~\ref{fig:InfiniteStrip=3-1.jpg}, and from $2.4$ to $3.75$ (shown
in fig~\ref{fig:InfiniteStrip=3-2.jpg}). The cases when $N=10$,
$N=20$, $N=30$ also exhibits similar behavior. They also have almost
two ranges of values for $m$ when two bands close at one point. Their
result are plotted are at fig~\ref{fig:InfiniteStrip=10.jpg},
fig~\ref{fig:InfiniteStrip=20.jpg}, and ... the result for $N=30$ will
be updated here when my program finishes running.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=3-1}.jpg}
    \caption{Infinite Strip, $N=3$, closes at $p_x=0$}
    \label{fig:InfiniteStrip=3-1.jpg}
\end{figure}
\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=3-2}.jpg}
    \caption{Infinite Strip, $N=3$, closes at $p_x=\pi$}
    \label{fig:InfiniteStrip=3-2.jpg}
\end{figure}
\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=10}.jpg}
    \caption{Infinite Strip, $N=10$}
    \label{fig:InfiniteStrip=10.jpg}
\end{figure}
\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=20}.jpg}
    \caption{Infinite Strip, $N=20$}
    \label{fig:InfiniteStrip=20.jpg}
\end{figure}

@@

    \subsubsection{Edge States Analytically Obtained}
    \label{sec:Edge States}
    I will analyse the continuum limit of this model at the point
    $(p_x=0,p_y=0)$, where the energy $E=0$. I also assume that we
    have placed to materials adjacent to each other in $x$ direction.
    I will assume that the two materials are in different topological
    states. For example, if the band cross when $m=0$, then one
    material will have $m<1$, and the other will have $m>0$.
    Therefore, at the interface, we could have $m$ as a function of
    $x$ such that $m(0)=0$.

@@

    \textbf{Remark}: Although I have broken translational symmetry in
    both direction, I could do almost the same calculation with $p_y$
    conserved, e.g. in a infinite strip. In that case, I would get a
    state localized in the interface, and have $p_y=-E$, representing
    an electron travelling with momentum $-E$ in $y$'s direction.
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\bibliography{../../Library}{}
\bibliographystyle{alphaurl}

>>>>

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physicist-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}
\usepackage{menukeys}
\makenomenclature % For an index of symbols.

@@

Calculations in this are contained in files: 
\begin{itemize}
    \item \menu{Lattice Dirac Model (2+1)-d-4-Infinite Stripe (N=3).nb},
    \item \menu{Lattice Dirac Model (2+1)-d-4-Infinite Stripe (N=20).nb},
    \item \menu{Lattice Dirac Model (2+1)-d-4-Infinite Stripe (N=30).nb},
\end{itemize}

This amounts to removing the $A$ in top right corner and $C$ in bottom left
corner of eq.\ref{eq:disc-eigeneq}.  In this case, the crossing bands exists at
slightly different value of $m$, also, there are few sharp transitions from a
closed gap state, to an open gap state. For example, when $N=3$, the band
crossing happens at two ranges of $m$: from $0$ to $2$ and from $2$ to $4$
(shown in Figure~\ref{fig:InfiniteStrip=3.jpg}). The cases when $N=10$, $N=20$,
$N=30$ also exhibits similar behavior. They also have almost two ranges of
values for $m$ when two bands close at one point. Their result are plotted are
at fig~\ref{fig:InfiniteStrip=10.jpg}, fig~\ref{fig:InfiniteStrip=20.jpg}, and
fig~\ref{fig:InfiniteStrip=30.jpg}.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=3}.jpg}
    \caption{Infinite Strip, $N=3$}
    \label{fig:InfiniteStrip=3.jpg}
\end{figure}
\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=10}.jpg}
    \caption{Infinite Strip, $N=10$}
    \label{fig:InfiniteStrip=10.jpg}
\end{figure}
\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=20}.jpg}
    \caption{Infinite Strip, $N=20$}
    \label{fig:InfiniteStrip=20.jpg}
\end{figure}
\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/InfiniteStrip/{InfiniteStrip.N=30}.jpg}
    \caption{Infinite Strip, $N=30$}
    \label{fig:InfiniteStrip=30.jpg}
\end{figure}

@@

    \subsubsection{Edge States in Continuum Limit}
    \label{sec:Edge States}
    I will analyse the continuum limit of this model at the point
    $(p_x=0,p_y=0)$, where the energy $E=0$. I also assume that we
    have placed to materials adjacent to each other in $x$ direction.
    I will assume that the two materials are in different topological
    states. For example, if the band cross when $m=0$, then one
    material will have $m<1$, and the other will have $m>0$.
    Therefore, at the interface, we could have $m$ as a function of
    $x$ such that $m(0)=0$.

@@

    \begin{remark}
    Although the above calculation breaks translational invariance in both
    direction, the actual boundary states breaks only one. Nonetheless, I could
    do almost the same calculation with $p_y$ conserved, e.g. in a infinite
    strip. In that case, I would get a state localized in the interface, and
    have $p_y=-E$, representing an electron travelling with momentum $-E$ in
    $y$'s direction.
    \end{remark}

    \subsubsection{Edge State in Infinity Strip}
    The edge state in this infinite strip (translational invariant in $x$, and
    open boundary in $y$) is difficult is obtain. The eigenvalue equation
    \ref{eq:disc-eigeneq} gives the following relation:
    \begin{equation}
        \label{eq:rec-disc-eigeneq}
        A \psi_{n-1} + B\psi_n + C\psi_{n+1} = E\psi_n
    \end{equation}
    If $A,B,C$ are numbers (or invertible matrices), then we could express
    $\psi_1$ in turns of $\psi_N$ (assuming $E=0$) and take the limit
    $\psi_1/\psi_N=0$ or $\infty$. This would hint a boundary state localized in
    $1$ or $N$.

    But now $A$ and $C$ are singular, this makes the
    equation~\ref{eq:rec-disc-eigeneq} mostly undetermined. The simplest case is
    that when $B=0$ (i.e. when $p_x=0$ or $\pm \pi$, and $2-m-\cos(p_x)=0$, thus
    $m=1$ or $3$),
    there are four possible eigenstates of eigenvalue $E=0$:
    \begin{align}
        &\psi_N = \begin{pmatrix}
            1 \\ -1
        \end{pmatrix}\text{ or } \begin{pmatrix}
            0 \\ 0
        \end{pmatrix},\,
        \psi_1 = \begin{pmatrix}
            1 \\ 1
        \end{pmatrix}\text{ or } \begin{pmatrix}
            0 \\ 0
        \end{pmatrix}\nonumber\\
        &\psi_{N-1}=\psi_{N-2} = \cdots = \psi_2 = \begin{pmatrix}
            0 \\ 0
        \end{pmatrix}
    \end{align}
    This shows that it possibly has states localized at the boundary, but this
    tells us nothing when $m$ is not the prescribed values($1$ and $3$).

    \subsection{Berry Phase}
    \label{sec:Berry Phase}
    The easiest case when we can calculate the Berry phase is when we have
    translational invariance in both $x$ and $y$ direction. The
    Fourier-transformed Hamiltonian is then diagonal. 
    \begin{align}
        &H =\sum_{p_x,p_y} c^\dagger_{p_x,p_y}\times
        \nonumber\\
        &\left[ \sin(p_x)\sigma^x + \sin(p_y)\sigma^y
            +(2-m-\cos(p_x)-\cos(p_y))\sigma^z
        \right] c_{p_x,p_y}
    \end{align}
    The Berry phase for a Hamiltonian of the form
    $\vec{d}(\vec{k})\vdot\vec{\sigma}$, is equal to $\pm\frac{1}{2}\Omega$,
    where the sign is determined by the choice of eigenstate, the $\Omega$ is
    the solid angle traversed by $\vec{d}$ as its parameter $\vec{k}$ varies,
    Now $\vec{d}=(\sin(p_x),\sin(p_y),2-m-\cos(p_x)-\cos(p_y))$, which is
    numerical explored\footnote{Details can be found in in file \menu{Attempts
    at Getting Chern Number.nb}.}.

    The surface of $\vec{d}$ as $p_x,p_y$ varies is somewhat distorted
    (fig~\ref{fig:BerryPhase/Shape}). Thankfully, it does not change as $m$
    varies (fig~\ref{fig:BerryPhase/AsMVaries}). Also, it is easy to see that
    when $m<0$ or $m>4$, the surface does not cut the $xy$ plane, hence in that
    region the Berry Phase is trivial. The distorted picture is dispatched into
    4 different pieces in fig~\ref{fig:BerryPhase/attempts}. It could be
    imagined that the vector $\vec{d}$ goes in the order: red
    $\to$blue$\to$yellow$\to$green$\to$red. With this and a little bit of
    imagination, I since the solid angle $\Omega$ when $0<m<2$ and $2<m<4$ are
    both $4\pi$.
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.5\linewidth]{pics/BerryPhase/Shape.pdf}
        \caption{Surface formed as $p_x,p_y$ varies}
        \label{fig:BerryPhase/Shape}
    \end{figure}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/BerryPhase/AsMVaries.pdf}
        \caption{As $m$ varies, the shape does not change. The blue plane is the
        $xy$ plane. The center of that surface meets the $xy$-plane when $m=2$.
        That surface leaves the $xy$-plane when $m<0$ or $m>4$.}
        \label{fig:BerryPhase/AsMVaries}
    \end{figure}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/BerryPhase/attempts.pdf}
        \caption{Attempts to determine the Berry phase by dispatching the surface
        into 4 different patches ($m=2$)}
        \label{fig:BerryPhase/attempts}
    \end{figure}
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\bibliography{../../Library}{}
\bibliographystyle{alphaurl}
